Global Distribution of Unintended Pregnancies and Contraceptive Usage

### 2024-04-15 code/eda-code/eda.qmd
Added labels to figures in manuscript, switched axis labels.

# Explore the percentage of unintended pregnancies related to short acting reversible methods in each country
p3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + 
  geom_histogram() + 
  ggtitle("Percent Unintended Pregnancy Using Short Acting Reversible Methods") + xlab("Percentage") + ylab("Number of Countries")
plot(p3)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png")
ggsave(filename = figure_file, plot=p3)

@@

P3: This histogram shows that the percentage of unintended pregnancies among those using short acting reversible methods varies more between countries compared to long acting reversible methods. About half of the countries in the dataset report less than 20% of pregnancies are unintended when women use this form of birth control.

@@

# Explore the total number of people using Short Acting Reversible Methods in each country
p7 <- mydata %>% ggplot(aes(x=total_sarc)) + 
  geom_histogram() + 
  ggtitle("Total Usage of Short Acting Reversible Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(p7)
figure_file = here("results", "figures", "exploratory-figures", "total_sarc_distribution.png")
ggsave(filename = figure_file, plot=p7)

@@

Plot 11 uses `pct_upreg_sarc` and `sarc_standard` to examine the same relationship as plot 10 but with short-acting reversible methods. 
```{r}

# explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
p11 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p11)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
ggsave(filename = figure_file, plot=p11)

@@

P11: The same positive trend is shown between the percentage of unintended pregnancies and number of women using SARMs, but there is considerably more variation in this relationship.

@@

# explore percent unintended pregnancies using short acting reversible methods by region
p15 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(p15)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
ggsave(filename = figure_file, plot=p15)

@@

P15: This graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.

@@

# explore short acting reversible method usage compared to percent of unintended pregnancies
p19 <- mydata %>% ggplot(aes(x = sarc_usage, y = pct_upreg_sarc)) +
  geom_violin() + ggtitle("Usage of SARCs vs Percent of Unintended Pregnancies") +
  labs(x = "Short Acting Reversible Usage", y = "Percentage Unintended Pregnancies") 
plot(p19)
figure_file = here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png")
ggsave(filename = figure_file, plot=p19)

@@

P19: It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.

We also need to explore the remaining predictor variables: `region`, `pct_currentlymarried`, `pct_nevermarried`, `rate_matdeaths`, `rate_matdeaths_upreg`, and `rate_matdeaths_abs`. Simple histograms will allow us to see the distribution of each variable and ensure there are no outliers. 
```{r}
# explore the region predictor
p22 <- mydata %>% ggplot(aes(x=region)) + 
  geom_bar() +
  ggtitle("Countries in Each Region") + 
  xlab("Region") + ylab("Number of Countries")
plot(p22)
figure_file = here("results", "figures", "exploratory-figures", "region_distribution.png")
ggsave(filename = figure_file, plot=p22)

# explore the current marriage predictor
p23 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
  geom_histogram() +
  ggtitle("Percent of Women Currently Married") + 
  xlab("Percentage") + ylab("Number of Countries")
plot(p23)
figure_file = here("results", "figures", "exploratory-figures", "currmarried_distribution.png")
ggsave(filename = figure_file, plot=p23)

# explore the never married predictor
p24 <- mydata %>% ggplot(aes(x=pct_nevermarried)) + 
  geom_histogram() +
  ggtitle("Percent of Women Never Married") + 
  xlab("Percentage") + ylab("Number of Countries")
plot(p24)
figure_file = here("results", "figures", "exploratory-figures", "nevmarried_distribution.png")
ggsave(filename = figure_file, plot=p24)

# explore the total maternal mortality rate predictor
p25 <- mydata %>% ggplot(aes(x=rate_matdeaths)) + 
  geom_histogram() +
  ggtitle("Total Rate of Maternal Deaths") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(p25)
figure_file = here("results", "figures", "exploratory-figures", "matdeaths_distribution.png")
ggsave(filename = figure_file, plot=p25)

@@

# explore the maternal mortality rate from unintended pregnancies predictor
p26 <- mydata %>% ggplot(aes(x=rate_matdeaths_upreg)) + 
  geom_histogram() +
  ggtitle("Rate of Maternal Deaths from Unintended Pregnancies") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(p26)
figure_file = here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png")
ggsave(filename = figure_file, plot=p26)

@@

# explore the maternal mortality rate from abortions
p27 <- mydata %>% ggplot(aes(x=rate_matdeaths_abs)) + 
  geom_histogram() +
  ggtitle("Rate of Maternal Deaths from Abortions") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(p27)
figure_file = here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png")
ggsave(filename = figure_file, plot=p27)

@@

```
P22: The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes. about 35 countries. Latin America/the Caribbean accounts for about 25 countries. Europe and Oceania account for the least number of observations.

P23: The distribution of women currently married is approximately normal with an average around 57% and no major outliers.

P24: The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%.

P25: The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers. We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality.

P26: The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram. Chad and South Sudan also account for the highest rates of maternal mortality in this category.

P27: The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor.

>>>>

# Explore the percentage of unintended pregnancies related to short acting reversible methods in each country
plot_1 <- mydata %>% ggplot(aes(x=pct_upreg_sarc)) + 
  geom_histogram() + 
  ggtitle("Percent Unintended Pregnancy Using Short Acting Reversible Methods") + xlab("Percentage") + ylab("Number of Countries")
plot(plot_1)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png")
ggsave(filename = figure_file, plot=plot_1)

@@

Plot 1: This histogram shows that the percentage of unintended pregnancies among those using short acting reversible methods varies more between countries compared to long acting reversible methods. About half of the countries in the dataset report less than 20% of pregnancies are unintended when women use this form of birth control.

@@

# Explore the total number of people using Short Acting Reversible Methods in each country
plot_2 <- mydata %>% ggplot(aes(x=total_sarc)) + 
  geom_histogram() + 
  ggtitle("Total Usage of Short Acting Reversible Methods per Country") + xlab("Number of Women") + ylab("Number of Countries")
plot(plot_2)
figure_file = here("results", "figures", "exploratory-figures", "total_sarc_distribution.png")
ggsave(filename = figure_file, plot=plot_2)

@@

Plot 3 uses `pct_upreg_sarc` and `sarc_standard` to examine the relationship with short-acting reversible methods. 
```{r}

# explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
plot_3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(plot_3)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
ggsave(filename = figure_file, plot=plot_3)

@@

Plot 3: The same positive trend is shown between the percentage of unintended pregnancies and number of women using SARMs, but there is considerably more variation in this relationship.

@@

# explore percent unintended pregnancies using short acting reversible methods by region
plot_4 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(plot_4)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
ggsave(filename = figure_file, plot=plot_4)

@@

Plot 4: This graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.

@@

# explore short acting reversible method usage compared to percent of unintended pregnancies
plot_5 <- mydata %>% ggplot(aes(x = sarc_usage, y = pct_upreg_sarc)) +
  geom_violin() + ggtitle("Usage of SARCs vs Percent of Unintended Pregnancies") +
  labs(x = "Short Acting Reversible Usage", y = "Percentage Unintended Pregnancies") 
plot(plot_5)
figure_file = here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png")
ggsave(filename = figure_file, plot=plot_5)

@@

Plot 5: It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method.

We also need to explore the remaining predictor variables: `region`, `pct_currentlymarried`, `pct_nevermarried`, `rate_matdeaths`, `rate_matdeaths_upreg`, and `rate_matdeaths_abs`. Simple histograms will allow us to see the distribution of each variable and ensure there are no outliers. 
```{r}
# explore the region predictor
plot_6 <- mydata %>% ggplot(aes(x=region)) + 
  geom_bar() +
  ggtitle("Countries in Each Region") + 
  xlab("Region") + ylab("Number of Countries")
plot(plot_6)
figure_file = here("results", "figures", "exploratory-figures", "region_distribution.png")
ggsave(filename = figure_file, plot=plot_6)

# explore the current marriage predictor
plot_7 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
  geom_histogram() +
  ggtitle("Percent of Women Currently Married") + 
  xlab("Percentage") + ylab("Number of Countries")
plot(plot_7)
figure_file = here("results", "figures", "exploratory-figures", "currmarried_distribution.png")
ggsave(filename = figure_file, plot=p23)

# explore the never married predictor
plot_8 <- mydata %>% ggplot(aes(x=pct_nevermarried)) + 
  geom_histogram() +
  ggtitle("Percent of Women Never Married") + 
  xlab("Percentage") + ylab("Number of Countries")
plot(plot_8)
figure_file = here("results", "figures", "exploratory-figures", "nevmarried_distribution.png")
ggsave(filename = figure_file, plot=plot_8)

# explore the total maternal mortality rate predictor
plot_9 <- mydata %>% ggplot(aes(x=rate_matdeaths)) + 
  geom_histogram() +
  ggtitle("Total Rate of Maternal Deaths") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(plot_9)
figure_file = here("results", "figures", "exploratory-figures", "matdeaths_distribution.png")
ggsave(filename = figure_file, plot=plot_9)

@@

# explore the maternal mortality rate from unintended pregnancies predictor
plot_10 <- mydata %>% ggplot(aes(x=rate_matdeaths_upreg)) + 
  geom_histogram() +
  ggtitle("Rate of Maternal Deaths from Unintended Pregnancies") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(plot_10)
figure_file = here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png")
ggsave(filename = figure_file, plot=plot_10)

@@

# explore the maternal mortality rate from abortions
plot_11 <- mydata %>% ggplot(aes(x=rate_matdeaths_abs)) + 
  geom_histogram() +
  ggtitle("Rate of Maternal Deaths from Abortions") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(plot_11)
figure_file = here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png")
ggsave(filename = figure_file, plot=plot_11)

@@

```
Plot 6: The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes. about 35 countries. Latin America/the Caribbean accounts for about 25 countries. Europe and Oceania account for the least number of observations.

Plot 7: The distribution of women currently married is approximately normal with an average around 57% and no major outliers. 
Plot 8: The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%.

Plot 9: The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers. We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality.

Plot 10: The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram. Chad and South Sudan also account for the highest rates of maternal mortality in this category.

Plot 11: The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor.

### 2024-04-15 products/manuscript/Manuscript.qmd
Added labels to figures in manuscript, switched axis labels.

After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen @tbl-summarytable.
```{r}
#| label: tbl-summarytable
#| tbl-cap: "Summary table of numeric variables"
#| echo: FALSE
predictortable<- readRDS(here("results", "tables", "predictortable.rds"))
predictortable
```

We created a multitude of exploratory plots to understand our data better. The first @exploratory-result shows the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods. The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
```{r}
#| label: exploratory-result
#| fig-cap: "Histograms of Percentage of Unintended Pregnancy Use by Birth Control Method"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
```

The next set of variables we explored were the total counts of women using each birth control method. There were two large outliers that prevented the graphs from being useful. After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years). There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected. This association is displayed in @exploratory-result2.
```{r}
#| label: exploratory-result2
#| fig-cap: "Relationship between Percentage of Unintended Pregnancies and Standardized Population Total"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
```

We stratified these four graphs by region to examine how the relationship differs across each of the t regions which is shown in @exploratory-result3 . The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
```{r}
#| label: exploratory-result3
#| fig-cap: "Percentage of Unintended Pregnancies Stratified by Region"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png"))
```

We also created violin plots using the categorical usage variables for each birth control method as shown in @exploratory-result4. It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method. The high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies. 
```{r}
#| label: exploratory-result4
#| fig-cap: "Violin Plots with Birth Control Usage Levels"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png"))
```

@@

```{r}
#| label: statistical-analysis1
#| fig-cap: "Strongest Predictor: sarc_standard"
#| echo: FALSE
readRDS(here("results","tables", "resulttable1.rds")) 
```

The second model we created used the region predictor, which revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region. The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association. These associations can be seen in @statistical-analysis2.
```{r}
#| label: statistical-analysis2
#| fig-cap: "Regional Predictor"
#| echo: FALSE
readRDS(here("results","tables", "resulttable2.rds")) 
```
Details about the 9 different models we created can be found in the statistical-analysis.qmd file in the code > analysis-code folder along with an exploration of the correlation between predictors.

@@

We wanted to analyze the differences in the outcome between regions further, so we completed an ANOVA test and subsequet Tukey HSD test. The Tukey HSD test for the region predictor, shown in @statistical-analysis3, revealed the largest differences in the outcome between the Oceania region and Latin America/the Caribbean, Europe, and Asia. The differences between Oceania and Europe and Latin America/the Caribbean are the largest. Oceania is clearly experiencing the lowest percentage of unintended pregnancies among short acting reversible method users. Two more significant differences in the outcome exist between Latin America/the Caribbean and Europe with Africa, which shows that Africa is experiencing higher levels of the outcome. 
```{r}
#| label: statistical-analysis3
#| fig-cap: "Tukey HSD Test Results"
#| echo: FALSE
readRDS(here("results","tables", "resulttable13.rds")) 
```
We created one model with all 5 predictors using multiple linear regression, shown in @statistcal-analysis4. This model shows the strongest predictor is the number of women using short acting reversible contraception. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. The currently married predictor is associated with a weak decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both show a very weak positive association with the outcome. 
```{r}
#| label: statistical-analysis4
#| fig-cap: "Multiple Linear Regression with 5 Predictors"
#| echo: FALSE
readRDS(here("results","tables", "resulttable8.rds")) 
```
To determine which model performed the best, we started by examining the RMSE and MAE performance metrics. We found that the model with the standardized total count of women using the short acting reversible contraception method performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion. @statistical-analysis5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors. The model with the population size using SARCs is the only single predictor model that performs better than the null. The full model with all 5 predictors peforms the best with the lowest RMSE and MAE values. 
```{r}
#| label: statistical-analysis5
#| fig-cap: "Performance Metric Comparisons"
#| echo: FALSE
readRDS(here("results","tables", "resulttable14.rds")) 
```
We used cross-validation of the two best models to determine how well the model performs on data it has not seen yet by splitting the data into testing and training pieces. The cross-validated results, shown in @statistical-analysis6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better. The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319. While the RMSE value for the multiple linear regression model is slightly higher than the first time it was calculated, it is still lower than the single predictor model. Additionally, the standard error of the RMSE for the multiple linear regression model is 0.2768, which is lower than the standard error of the RMSE for the single predictor model of 0.3528. This means there is higher variance in the single predictor model. The R-squared value of 0.8206 for the multiple linear regression model is higher than 0.7861 for the single predictor model, which means the multiple linear regression model accounts for more of the variability in the data. Overall, the multiple linear regression model performs better.
```{r}
#| label: statistical-analysis6
#| fig-cap: "Cross Validation Performance Metric Comparisons"
#| echo: FALSE
readRDS(here("results","tables", "resulttable15.rds")) 
```

Since we determined that the full model with all 5 predictors performs best, we graphed the predicted versus observed values using this model with both the testing and training data to further examine model performance. As seen in @statistical-analysis7, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line. The testing model predictions, shown in coral, are much more extreme than the training model predictions because they are farther from the 45 degree line. While there is some overlap between the two sets of predictions, it is likely that our model is overfitting the data. 
```{r}
#| label: statistical-analysis7
#| fig-cap: "Evaluating Final Model Performance on Test Data"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "fullmodelpred.png"))
```
We went one step further to examine the uncertainty in the full model predictions using the bootstrap method with 100 resamples of the training data. The figure shown in @statistical-analysis8 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model. The confidence intervals are narrow, which shows there is minimal uncertainty in the model prediction. For the predictions that do not fall on the expected 45 degree line, the confidence intervals do not intersect with the line either. These findings increase the validity of the conclusion that the model is overfitting the data because there is not much uncertainty in the predictions. 
```{r}
#| label: statistical-analysis8
#| fig-cap: "Evaluation of Uncertainty in Full Model Predictions"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "bootstrappred.png"))
```

Since we have evidence that our model overfits the data, the last technique we want to use to explore our full model is Lasso regression. This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting. As seen in @statistical-analysis9, the LASSO regression coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model. The `sarc_standard` predictor still has the strongest relationship with the outcome. The percent of women currently married, rate of maternal deaths due to unintended pregnancies, and rate of maternal deaths to due to abortion all have weak negative associations with the outcome. This process confirms that our final model should contain all 5 predictors. 
```{r}
#| label: statistical-analysis9
#| fig-cap: "LASSO Regression Results"
#| echo: FALSE
readRDS(here("results","tables", "resulttable16.rds")) 
```
It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @statistical-analysis10. As the regularization penalty increases past 1, we see that the RMSE value increases exponentially. The lowest RMSE for the model is slightly greater than 5 when the regularization penalty is less than 1. When the regularization penalty increases to 10, the RMSE increases to almost 11. Since we originally chose 0.1 as our penalty parameter, we found one of the best performing models with the LASSO regression based on an RMSE value slightly greater than 5, which i comparable to the RMSE found in the first regression model with all 5 predictors. 
```{r}
#| label: statistical-analysis10
#| fig-cap: "Parameter Tuning for LASSO Regression"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "lassoparameters.png"))
```

>>>>

After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen @tbl-summarytable.
```{r}
#| label: tbl-summarytable
#| tbl-cap: "Table 1: Summary Table Of Numeric Variables"
#| echo: FALSE
predictortable<- readRDS(here("results", "tables", "predictortable.rds"))
predictortable
```

We created a multitude of exploratory plots to understand our data better. The first @exploratory-result shows the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods. The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
```{r}
#| label: exploratory-result
#| fig-cap: "Plot 1: Histogram Depicting Number of Countries With Percent Unintended Pregnancy Using SARC"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
```

The next set of variables we explored were the total counts of women using each birth control method. There were two large outliers that prevented the graphs from being useful. After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years). There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected. This association is displayed in @exploratory-result2.
```{r}
#| label: exploratory-result2
#| fig-cap: "Figure 1: Graph Depciting Percent Unintended Pregnancy by Standarized Count of Total Women"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
```

We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @exploratory-result3 . The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
```{r}
#| label: exploratory-result3
#| fig-cap: "Figure 2: Graph Depicting Percent Unintended Pregnancies Using SARC by region"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png"))
```

We also created violin plots using the categorical usage variables for each birth control method as shown in @exploratory-result4. It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method. The high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies. 
```{r}
#| label: exploratory-result4
#| fig-cap: "Plot 5: Violin Plot Depicitng Usage of SARc Per Region"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png"))
```

@@

```{r}
#| label: statistical-analysis1
#| fig-cap: "Table 2: Strongest Predictor: sarc_standard"
#| echo: FALSE
readRDS(here("results","tables", "resulttable1.rds")) 
```

The second model we created used the region predictor, which revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region. The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association. These associations can be seen in @statistical-analysis2.
```{r}
#| label: statistical-analysis2
#| fig-cap: "Table 3: Regional Predictor"
#| echo: FALSE
readRDS(here("results","tables", "resulttable2.rds")) 
```
Details about the 9 different models we created can be found in the statistical-analysis.qmd file in the code > analysis-code folder along with an exploration of the correlation between predictors.

@@

We created one model with all 5 predictors using multiple linear regression, shown in @statistcal-analysis4. This model shows the strongest predictor is the number of women using short acting reversible contraception. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. The currently married predictor is associated with a weak decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both show a very weak positive association with the outcome. 
```{r}
#| label: statistical-analysis4
#| fig-cap: "Table 4: Multiple Linear Regression with 5 Predictors"
#| echo: FALSE
readRDS(here("results","tables", "resulttable8.rds")) 
```
To determine which model performed the best, we started by examining the RMSE and MAE performance metrics. We found that the model with the standardized total count of women using the short acting reversible contraception method performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion. @statistical-analysis5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors. The model with the population size using SARCs is the only single predictor model that performs better than the null. The full model with all 5 predictors peforms the best with the lowest RMSE and MAE values. 
```{r}
#| label: statistical-analysis5
#| fig-cap: "Table 5: Performance Metric Comparisons"
#| echo: FALSE
readRDS(here("results","tables", "resulttable14.rds")) 
```
We used cross-validation of the two best models to determine how well the model performs on data it has not seen yet by splitting the data into testing and training pieces. The cross-validated results, shown in @statistical-analysis6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better. The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319. While the RMSE value for the multiple linear regression model is slightly higher than the first time it was calculated, it is still lower than the single predictor model. Additionally, the standard error of the RMSE for the multiple linear regression model is 0.2768, which is lower than the standard error of the RMSE for the single predictor model of 0.3528. This means there is higher variance in the single predictor model. The R-squared value of 0.8206 for the multiple linear regression model is higher than 0.7861 for the single predictor model, which means the multiple linear regression model accounts for more of the variability in the data. Overall, the multiple linear regression model performs better.
```{r}
#| label: statistical-analysis6
#| fig-cap: "Table 6: Cross Validation Performance Metric Comparisons"
#| echo: FALSE
readRDS(here("results","tables", "resulttable15.rds")) 
```

Since we determined that the full model with all 5 predictors performs best, we graphed the predicted versus observed values using this model with both the testing and training data to further examine model performance. As seen in @statistical-analysis7, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line. The testing model predictions, shown in coral, are much more extreme than the training model predictions because they are farther from the 45 degree line. While there is some overlap between the two sets of predictions, it is likely that our model is overfitting the data. 
```{r}
#| label: statistical-analysis7
#| fig-cap: "Figure 3: Evaluating Final Model Performance on Test Data"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "fullmodelpred.png"))
```
We went one step further to examine the uncertainty in the full model predictions using the bootstrap method with 100 resamples of the training data. The figure shown in @statistical-analysis8 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model. The confidence intervals are narrow, which shows there is minimal uncertainty in the model prediction. For the predictions that do not fall on the expected 45 degree line, the confidence intervals do not intersect with the line either. These findings increase the validity of the conclusion that the model is overfitting the data because there is not much uncertainty in the predictions. 
```{r}
#| label: statistical-analysis8
#| fig-cap: "Figure 4: Evaluation of Uncertainty in Full Model Predictions"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "bootstrappred.png"))
```

Since we have evidence that our model overfits the data, the last technique we want to use to explore our full model is Lasso regression. This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting. As seen in @statistical-analysis9, the LASSO regression coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model. The `sarc_standard` predictor still has the strongest relationship with the outcome. The percent of women currently married, rate of maternal deaths due to unintended pregnancies, and rate of maternal deaths to due to abortion all have weak negative associations with the outcome. This process confirms that our final model should contain all 5 predictors. 
```{r}
#| label: statistical-analysis9
#| fig-cap: "Table 7: LASSO Regression Results"
#| echo: FALSE
readRDS(here("results","tables", "resulttable16.rds")) 
```
It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @statistical-analysis10. As the regularization penalty increases past 1, we see that the RMSE value increases exponentially. The lowest RMSE for the model is slightly greater than 5 when the regularization penalty is less than 1. When the regularization penalty increases to 10, the RMSE increases to almost 11. Since we originally chose 0.1 as our penalty parameter, we found one of the best performing models with the LASSO regression based on an RMSE value slightly greater than 5, which i comparable to the RMSE found in the first regression model with all 5 predictors. 
```{r}
#| label: statistical-analysis10
#| fig-cap: "Figure 5: Parameter Tuning for LASSO Regression"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "lassoparameters.png"))
```
